Naturalness in U(1) extensions of the MSSM

### 2014-11-11 paper/e6ssm-tuning-paper.tex
fixed c+p errors in USSM superpotential.
fixed typo extar --> extra.
WITH --> with.

\begin{abstract}
% Copied SUSY2014 abstract as placeholder
With the discovery of a 125 GeV Standard Model (SM) like Higgs boson
and rising lower bounds on the masses of superpartners, minimal models
of low energy supersymmetry (SUSY) appear to be increasingly fine
tuned. In the Minimal Supersymmetric Standard Model (MSSM) large stop
masses, needed to obtain the observed Higgs mass, give a significant
contribution to fine tuning. $U(1)$ extensions of the MSSM that solve
the $\mu$-problem have new F- and D-term contributions to the Higgs mass,
allowing a 125 GeV Higgs with lighter stops and so alleviating the
associated fine tuning. However, it was recently demonstrated, within
a constrained version of the Exceptional Supersymmetric Standard Model (E$_6$SSM)
defined at the Grand Unification (GUT) scale, that the presence of a massive $Z'$ boson introduces a 
substantial new source of tuning in these theories at tree level. Here we
consider the impact of this tuning when we remove the fine tuning that depends
on assumptions about the SUSY breaking scale, by setting the SUSY
breaking parameters at low energies. We find that the existing
experimental bounds on the $Z'$ mass impose an effective lower bound on
the fine tuning, which does not depend on assumptions about SUSY breaking. 
We compare our results against the tuning in the MSSM WITH the SUSY breaking terms also defined at low energies.

@@

  The couplings for the $U(1)^\prime$ gauge group should allow the following renormalisable superpotential terms required in the USSM,
\be
W_{USSM} =  y^U_{ijk} \hat{u}^c_i \hat{H}^u_{j} \hat{Q}_k + y^D_{ijk} \hat{d}^c_i \hat{H}^d_{j} \hat{Q}_k + y^E_{ijk} \hat{e}^c_i \hat{H}^d_{j} \hat{L}_k + \lambda \hat{S} \hat{H}^d \hat{H},
\ee, with $i,j,k \in \{1,2,3\}$.

@@

The breaking of $E_6$ into $SO(10)$, gives rise to $E_6\to
SO(10)\times U(1)_{\psi}$, and the subsequent breaking of $SO(10)$
into $SU(5)$, gives $SO(10)\to SU(5)\times U(1)_{\chi}$ (this is
reviewed in e.g.~\cite{Langacker:2008yv})). The extar $U(1)$ gauge symmetry at low energies should then be a linear combination of these in the $E_6$ inspired case,
\be
U(1)^\prime = \cos(\theta) U(1)_{\chi} + \sin(\theta) U(1)_{\psi}.
\ee In table \ref{tab:E6charges} the charges for several popular $E_6$ inspired $U(1)$ extensions are shown.
% Table of U(1)' charges, which may or may not be useful to include.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
\hline
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\end{tabular}
\end{ruledtabular}
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta_{E_6}=\arctan\sqrt{15}$.}
\label{tab:E6charges}
\end{table}

>>>>

\begin{abstract}
% Copied SUSY2014 abstract as placeholder
With the discovery of a 125 GeV Standard Model (SM) like Higgs boson
and rising lower bounds on the masses of superpartners, minimal models
of low energy supersymmetry (SUSY) appear to be increasingly fine
tuned. In the Minimal Supersymmetric Standard Model (MSSM) large stop
masses, needed to obtain the observed Higgs mass, give a significant
contribution to fine tuning. $U(1)$ extensions of the MSSM that solve
the $\mu$-problem have new F- and D-term contributions to the Higgs mass,
allowing a 125 GeV Higgs with lighter stops and so alleviating the
associated fine tuning. However, it was recently demonstrated, within
a constrained version of the Exceptional Supersymmetric Standard Model (E$_6$SSM)
defined at the Grand Unification (GUT) scale, that the presence of a massive $Z'$ boson introduces a 
substantial new source of tuning in these theories at tree level. Here we
consider the impact of this tuning when we remove the fine tuning that depends
on assumptions about the SUSY breaking scale, by setting the SUSY
breaking parameters at low energies. We find that the existing
experimental bounds on the $Z'$ mass impose an effective lower bound on
the fine tuning, which does not depend on assumptions about SUSY breaking. 
We compare our results against the tuning in the MSSM with the SUSY breaking terms also defined at low energies.

@@

  The couplings for the $U(1)^\prime$ gauge group should allow the following renormalisable superpotential terms required in the USSM,
\be
W_{USSM} =  y^U_{ij} \hat{u}^c_i \hat{H}_u \hat{Q}_j + y^D_{ij} \hat{d}^c_i \hat{H}_d \hat{Q}_j + y^E_{ij} \hat{e}^c_i \hat{H}_d \hat{L}_j + \lambda \hat{S} \hat{H}_d \hat{H}_u,
\ee with $i,j \in \{1,2,3\}$.

@@

The breaking of $E_6$ into $SO(10)$, gives rise to $E_6\to
SO(10)\times U(1)_{\psi}$, and the subsequent breaking of $SO(10)$
into $SU(5)$, gives $SO(10)\to SU(5)\times U(1)_{\chi}$ (this is
reviewed in e.g.~\cite{Langacker:2008yv})). The extra $U(1)$ gauge symmetry at low energies should then be a linear combination of these in the $E_6$ inspired case,
\be
U(1)^\prime = \cos(\theta) U(1)_{\chi} + \sin(\theta) U(1)_{\psi}.
\ee In table \ref{tab:E6charges} the charges for several popular $E_6$ inspired $U(1)$ extensions are shown.
% Table of U(1)' charges, which may or may not be useful to include.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
\hline
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\end{tabular}
\end{ruledtabular}
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta_{E_6}=\arctan\sqrt{15}$.}
\label{tab:E6charges}
\end{table}
